### 2024-09-13 outputs/paper/paper.qmd
Add data part and 2 results part

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Junbo Li
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Package setup ####

library(tidyverse)
library(here)
library(tinytex)
library(dplyr)
library(tibble)
library(opendatatoronto)
library(janitor)
library(ggplot2)
library(knitr)

# Read data
raw_lead_data <- read_csv(here("inputs/data/raw_data.csv"))
cleaned_data <- read_csv(here("outputs/data/analysis_data.csv"))
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Raw data

The data used in here is downloaded from opendattoronto [@citeOpenDataToronto]. Specifically, we used the Shelter System Flow Data
released by Toronto government [@toronto_shelter_data].The data loading, cleaning and analysis process is done by R [@citeR], togehter with packges include tidyverse [@tidy]... This data measures number of people requiring for shelter inside city of Toronto from Jan-2018 till Jun-2024 by each month (actively_homeless). Also, this dataset provides division of age group, refugee status, gender, as well as chronic status to provide further detailed study into the change of composition of homeless people over time. While there are several other dataset including daily shelter occupacancy in opendatatoronto, they do not have the subdivision information such as age or refugee status as this one.

## Cleaned Data

We dropeed the only missing row corresponding to july-2024. We selected our columns of interest, namely, gender: ('male', 'female', 'transgender_non_binary_or_two_spirit'), total number of homeless people every month ('actively_homeless'), as well as refugee and chronic status in polulation column. In addition, we grouped age group by whether they are suitable for work (age under 24 for early career stage, age25-44 for working age and age above 45 for pre-retirement and retirement age). Finally, we replace the month data by averaged quatered data for better analysis and visualization of a prelonged time period. The cleaned data could be seen in @tbl-cleandata.

```{r}
#| label: tbl-cleandata
#| echo: false
#| tbl-cap: "cleaned dataset"

cleaned_data |>
  slice(1:5) |>
  kable(
    col.names = c("Time (Quaters)", "Subdivision", "Total people", "Female", "Male", "Other Gender",
                  "Early Career Stage", "Working Age", "Pre-retire and Retire"),
    digits = 0)

```

# Results {#sec-results}

## Overall trend

First, we plot the bar graph of over all trend of homeless people over the pass of the time. In @fig-overall, we observe an slightly increase of total number of homeless people from first quarter 2018 to second quarter of 2024. There is a decrease of homeless population in late 2020-early 2021, but increased quickly after that.

```{r}
#| label: fig-overall
#| echo: false
#| fig-cap: "Homeless Population Over Quarters"

# Filter the data for 'All Population' group
all_population_data <- subset(cleaned_data, population_group == 'All Population')

# Create the bar graph
ggplot(all_population_data, aes(x = Quarter, y = actively_homeless)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Quarters of a year",
       y = "Total number of homeless people") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
#| label: fig-gender
#| echo: false
#| fig-cap: "Proportion of Active Homeless Population by Gender Over Quarters"

# Filter the data for 'All Population' group
all_population_data <- subset(cleaned_data, population_group == 'All Population')

# Calculate the proportion of each gender
all_population_data <- all_population_data %>%
  mutate(total = gender_female + gender_male + gender_transgender_non_binary_or_two_spirit,
         female_proportion = gender_female / total,
         male_proportion = gender_male / total,
         transgender_proportion = gender_transgender_non_binary_or_two_spirit / total)

# Reshape the data for plotting
proportion_data_long <- all_population_data %>%
  select(Quarter, female_proportion, male_proportion, transgender_proportion) %>%
  gather(key = "Gender", value = "Proportion", -Quarter)

# Rename gender categories for better readability
proportion_data_long$Gender <- recode(proportion_data_long$Gender,
                                      female_proportion = "Female",
                                      male_proportion = "Male",
                                      transgender_proportion = "Transgender/Non-binary/Two-spirit")

# Create the stacked bar graph
ggplot(proportion_data_long, aes(x = Quarter, y = Proportion, fill = Gender)) +
  geom_bar(stat = "identity") +
  labs(x = "Quarters of a year",
       y = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("Female" = "pink", "Male" = "blue", "Other" = "purple"))

```

## Gender trend
Next, we investigate the trend of gender proportion. We plot the proportion of male, female, and Other genders against time in @fig-gender. Even thought there is some flatuations of the proportion of these three gender types, the overall trend is quite stable. The male proportion is around 60% while female proportion around 40% and others around 1%.

## Chronic trend

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References

### 2024-09-14 outputs/paper/paper.qmd
Finished result part

@@
                                       male_proportion = "Male",
-                                      transgender_proportion = "Transgender/Non-binary/Two-spirit")
+                                      transgender_proportion = "Other")
 
@@
 ## Chronic trend
+@fig-Chronic shows the trend of the proportion of chronic population among all of homeless people over the quarters of a year.
+Here, according to @chronic, chronic homeless refers to, "People who meet one of the two following criteria, as per the federal definition of chronic homelessness.
+The person has recorded a minimum of 180 overnight stay in the past year (365 days); or the person has recurrent overnight stays over the past three years with a cumulative duration of at least 546 nights." We can see form the graph that the proportion in chronic homeless increased a lot, from less than 35% in first quarter of 2018 to more than 65% in the second quarter of 2024.
+```{r}
+#| label: fig-Chronic
+#| echo: false
+#| fig-cap: "Proportion of Chronic population to Active Population by Quarter"
+
+# Filter the data for 'All Population' and 'Chronic'
+all_population <- subset(cleaned_data, population_group == "All Population")
+chronic_population <- subset(cleaned_data, population_group == "Chronic")
+
+# Calculate the proportion of chronic to active population
+proportion <- chronic_population$actively_homeless / all_population$actively_homeless
+
+# Create a data frame for plotting
+plot_data <- data.frame(Quarter = all_population$Quarter, Proportion = proportion)
+
+# Create a bar graph
+ggplot(plot_data, aes(x = Quarter, y = Proportion)) +
+  geom_bar(stat = "identity", fill = "blue") +
+  labs(x = "Quarters", y = "Proportion of Chronic to Active Population") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+## Age trend
+@fig-WorkingAge shows the trend of the proportion of different working age statuses among the actively homeless population over the quarters of a year.
+Here, the working age statuses are categorized into three groups: Early Career Stage, Working Age, and Pre-Retirement and Retirement Age.Even thought the trend is not huge, We can see from the graph that the proportion of the Early Career Stage population  as well as pre-retire and retire have decreased (from 29% and 33% separately in first quarter of 2018 to 25% and 28% in second quarter in 2024).
+On the other hand, working age proportion increased to 47% in 2024Q2 from 37% in 2018Q1.
+```{r}
+#| label: fig-WorkingAge
+#| echo: false
+#| fig-cap: "Proportion of working age population to Active Population by Quarter"
+
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+
+# Filter the data for 'All Population' group
+all_population_data <- subset(cleaned_data, population_group == 'All Population')
+
+# Calculate the proportion of each working age status
+all_population_data <- all_population_data %>%
+  mutate(total = Early_Career_Stage + Working_Age + Pre_Retirement_and_Retirement_Age,
+         early_career_proportion = Early_Career_Stage / total,
+         working_age_proportion = Working_Age / total,
+         pre_retirement_proportion = Pre_Retirement_and_Retirement_Age / total)
+
+# Reshape the data for plotting
+proportion_data_long <- all_population_data %>%
+  select(Quarter, early_career_proportion, working_age_proportion, pre_retirement_proportion) %>%
+  gather(key = "Age_Status", value = "Proportion", -Quarter)
+
+# Rename age status categories for better readability
+proportion_data_long$Age_Status <- recode(proportion_data_long$Age_Status,
+                                          early_career_proportion = "Early Career Stage",
+                                          working_age_proportion = "Working Age",
+                                          pre_retirement_proportion = "Pre-Retirement and Retirement Age")
+
+# Create the stacked bar graph
+ggplot(proportion_data_long, aes(x = Quarter, y = Proportion, fill = Age_Status)) +
+  geom_bar(stat = "identity") +
+  labs(x = "Quarters of a year",
+       y = "Proportion") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(legend.position = "bottom") +
+  scale_fill_manual(values = c("Early Career Stage" = "blue", 
+                               "Working Age" = "pink", 
+                               "Pre-Retirement and Retirement Age" = "green"))
+```
+
+## Refugee trend
+@fig-refugee shows the trend of the proportion of refugees among the actively homeless population over the quarters of a year.
+We observe a signifiacant decrease of the trend accompanyed by even more signifcant increase.
+```{r}
+#| label: fig-refugee
+#| echo: false
+#| fig-cap: "Proportion of Chronic population to Active Population by Quarter"
+
+# Filter the data for 'All Population' and 'Chronic'
+all_population <- subset(cleaned_data, population_group == "All Population")
+refu_population <- subset(cleaned_data, population_group == "Refugees")
+
+# Calculate the proportion of chronic to active population
+proportion <- refu_population$actively_homeless / all_population$actively_homeless
+
+# Create a data frame for plotting
+plot_data <- data.frame(Quarter = all_population$Quarter, Proportion = proportion)
+
+# Create a bar graph
+ggplot(plot_data, aes(x = Quarter, y = Proportion)) +
+  geom_bar(stat = "identity", fill = "blue") +
+  labs(x = "Quarters", y = "Proportion of Refugee to Active Population") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 

### 2024-09-14 outputs/paper/paper.qmd
Finished discussion part

@@
 # Discussion
+Based on the analysis of results section, we observe that, even though the overall population of homeless people did not incur sgnificant variation in the past six years, its demographics changed a lot.
+Especially, the proportion of refugees, chronic people as well as working aged people.
+In the past, City of Toronto has devolted its effort on reducing the number of homeless people, espically through projects of building affordable houses @torontosol.
+However, our study show that with the change of demographic, the city government should also give its focus on other strategies of reducing homelessness.
+
+First, as the result section points out, the proportion of chronic homeless people nearly doubled in the past six years.
+Despite its nature of long lasting homeless status, studies have also shown that chronic homelessness is closed associated with chronic illness or mental problems @chronic.
+As point by another study, average income of people experiencing mental illness and other types of disorders is significantly lower than normal people @Sareen2011.
+Thus, it is very hard for those people who have already experiencing mental and physical illness to escape homeless.
+In addition, homeless people would face more severe health problem due to lack of adequate food, discrimination, barriers to health care etc @Sleet2021.
+A combination of these would become an obvious negative cycle: illness lead to decrease of income, which leads to become homeless, which would further depend illness.
+What the city government can do, on the other hand, is to increase public health care system, espically focused to those low-income and homeless people.
+In fact, merely building houses only increased the proportion of chronic homeless people besides its aim is to reduce @house_not.
 

### 2024-09-14 outputs/paper/paper.qmd
finishes discussion

--- 2024-09-14 outputs/paper/paper.qmd
@@
 # Discussion
+Based on the analysis in the results section, we observe that although the overall population of homeless individuals in Toronto has not changed significantly over the past six years, its demographics have shifted considerably.
+Specifically, the proportions of refugees, chronically homeless individuals, and working-aged people have increased.
+In the past, the City of Toronto focused its efforts on reducing homelessness, primarily through projects aimed at building affordable housing @torontosol.
+However, our study indicates that with these demographic changes, the city government must also adopt additional strategies to address homelessness effectively.
+
+First, as highlighted in the results section, the proportion of chronically homeless individuals has nearly doubled in the past six years.
+While chronic homelessness inherently involves long-lasting periods without stable housing, studies have shown that it is closely associated with chronic illness and mental health issues @chronic.
+Research further reveals that individuals experiencing mental illness or other disorders tend to have significantly lower incomes than the general population @Sareen2011.
+Consequently, those already facing mental and physical health challenges find it exceptionally difficult to escape homelessness.
+Additionally, homeless individuals often encounter more severe health problems due to inadequate nutrition, discrimination, and barriers to healthcare @Sleet2021.
+These factors create a vicious cycle: illness leads to decreased income, which leads to homelessness, which in turn worsens the illness.
+
+To break this cycle, the city government should strengthen public healthcare systems, particularly for low-income and homeless populations.
+Building affordable housing alone has, in some cases, led to an increase in the proportion of chronically homeless individuals @house_not.
+Instead, the city should invest in supportive housing, where healthcare, mental health services, and social support are integrated with affordable housing.
+Given the strong link between chronic homelessness and mental health challenges, the city should also expand access to mental health and addiction services.
+Mobile outreach teams and community-based mental health clinics can help ensure that homeless individuals receive the care they need.
+Programs that reduce healthcare barriers—such as mobile health clinics, telehealth services, and partnerships with local hospitals—will improve health outcomes for homeless individuals and prevent them from becoming chronically homeless.
+
+Second, due to increased global conflicts in Eastern Europe and the Middle East, the number of refugees entering Toronto has risen significantly @cbc23.
+However, the city has been unprepared to support this influx, as the proportion of refugees among the homeless population has surged from less than 20% in the first three quarters of 2021 to nearly 60% by the second quarter of 2024.
+There is an urgent need for the government to accelerate its efforts in assisting newly arrived refugees, providing them with jobs and secure housing.
+
+Finally, we observe an increase in the proportion of working-aged homeless individuals.
+This underscores the need for the government to create more job opportunities, particularly for those who rely on employment to make a living, such as individuals experiencing mental health challenges or physical disabilities.
 

### 2024-09-14 outputs/paper/paper.qmd
First edit of discussion

--- 2024-09-14 outputs/paper/paper.qmd
@@
 # Discussion
+Based on the analysis of results section, we observe that, even though the overall population of homeless people did not incur sgnificant variation in the past six years, its demographics changed a lot.
+Especially, the proportion of refugees, chronic people as well as working aged people.
+In the past, City of Toronto has devolted its effort on reducing the number of homeless people, espically through projects of building affordable houses @torontosol.
+However, our study show that with the change of demographic, the city government should also give its focus on other strategies of reducing homelessness.
+
+First, as the result section points out, the proportion of chronically homeless people nearly doubled in the past six years.
+Despite its nature of long-lasting homeless status, studies have also shown that chronic homelessness is closely associated with chronic illness or mental problems @chronic.
+As pointed out by another study, the average income of people experiencing mental illness and other types of disorders is significantly lower than that of normal people @Sareen2011.
+Thus, it is very hard for those people who have already experienced mental and physical illness to escape homelessness.
+In addition, homeless people would face more severe health problems due to lack of adequate food, discrimination, barriers to health care, etc @Sleet2021.
+A combination of these would become an obvious negative cycle: illness leads to a decrease in income, which leads to becoming homeless, which would further depend on illness.
+What the city government can do, on the other hand, is to increase the public health care system, especially focused on low-income and homeless people.
+In fact, merely building houses only increased the proportion of chronically homeless people and its aim is to reduce @house_not.
+
+Second, due to increased global conflict in Eastern Europe and the Middle East, the number of refugees entering Toronto has increased @cbc23.On the other hand, the Toronto government has not been prepared to support such a huge increase of refugees as the portion of refugees in homeless people roared from less than 20% in the first three quarters of 2021 to nearly 60% at second quarter 2024.
+We observe an emergency for the government to increase its speed of helping these newly entered refugees and give them a job and a secure place to live.
+
+Finally, we observe an increase in the proportion of people of worked-age.
+This indicates the need of the government to create more jobs for the people who need it to make a living, like people who are experiencing mental health problems or physical disabilities.
 

### 2024-09-14 outputs/paper/paper.qmd
Intro and Liminations added

@@
 format: pdf
+toc: true
 number-sections: true
@@
 # Introduction
+According to @canadianobservatory2012definition, homeless is defined as "the situation of an individual, family or community without stable, safe, permanent, appropriate housing, or the immediate prospect, means and ability of acquiring it".
+It is estimated that about 23,500 Canadian people experienced every type of Homeless every year @gaetz2014state.
+On SEP 12, 2024, there are still more than 9500 people staying in shelters in the city of Toronto @toronto2024shelter.
+To reduce homelessness, Toronto city government published the supportive houseing development plan, which aims to build more affordable housings for the homeless people @toronto2024affordable.
+A recent two-year plan starting 2022 aiming to build 4,000 more affordable homes as well as supportive houseings by the end of 2024.
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+In this paper, we studied the demographical shift of total homeless populations in the city of Toronto, and find that the underlyin demographics of homesless populations changed significantly.
+This would indicate that the tradition plan which mainly focus on buliding houses would no longer solve new homeless problem.
 
-The remainder of this paper is structured as follows. @sec-data....
+In the following sections, we will first introduce our sourse of data as well as potential cleaning procedures.
+Then, we will show our results of the demographic change.
+Finally, we would indicate the reason of our claim base on the analysis of the result and give some suggestions that would help better reduce homelessness in the city.
 
@@
 # Discussion
-Based on the analysis of results section, we observe that, even though the overall population of homeless people did not incur sgnificant variation in the past six years, its demographics changed a lot.
-Especially, the proportion of refugees, chronic people as well as working aged people.
-In the past, City of Toronto has devolted its effort on reducing the number of homeless people, espically through projects of building affordable houses @torontosol.
-However, our study show that with the change of demographic, the city government should also give its focus on other strategies of reducing homelessness.
+Based on the analysis of the results section, we observe that, even though the overall population of homeless people did not incur significant variation in the past six years, its demographics changed a lot.
+Specifically, the proportion of refugees, chronic people as well as working-aged people.
+In the past, the City of Toronto has devoted its efforts to reducing the number of homeless people, especially through projects of building affordable houses [@torontosol].
+However, our study shows that with the change of demographics, the city government should also give its focus on other strategies for reducing homelessness.
 
@@
 Thus, it is very hard for those people who have already experienced mental and physical illness to escape homelessness.
-In addition, homeless people would face more severe health problems due to lack of adequate food, discrimination, barriers to health care, etc @Sleet2021.
+In addition, homeless people would face more severe health problems due to lack of adequate food, discrimination, barriers to health care, etc [@Sleet2021].
 A combination of these would become an obvious negative cycle: illness leads to a decrease in income, which leads to becoming homeless, which would further depend on illness.
 What the city government can do, on the other hand, is to increase the public health care system, especially focused on low-income and homeless people.
-In fact, merely building houses only increased the proportion of chronically homeless people and its aim is to reduce @house_not.
+In fact, merely building houses only increased the proportion of chronically homeless people and its aim is to reduce [@house_not].
 
@@
 
-Finally, we observe an increase in the proportion of people of worked-age.
-This indicates the need of the government to create more jobs for the people who need it to make a living, like people who are experiencing mental health problems or physical disabilities.
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
+Finally, we observe an increase in the proportion of people of working age.
+This indicates the need for the government to create more jobs for the people who need them to make a living, like people who are experiencing mental health problems or physical disabilities.
 
@@
 
-Weaknesses and next steps should also be included.
+Even though our data is the most uptodate and accurate date aviliable for homeless population on opendatatoronto, it might have potential bias towards estimation the ture homeless population in city of Toronto.
+One clear bias is that people would not stay in the shelter everyday and they would enter and exit the shelter frequently [@toronto_shelter_data].
+Correctly measure the ture popution demegraphic is a very diffcult on-going problem in Canada, but the shelter flow data would provide a proper estimate of our interest [@dionne2023review].
+To understand the probelms better and develop more strategies the next steps would be to research stratgies that could reach out and survey the population that is diffcult to contact such as unsheltered homeless population.
+Also, a new homeless reduction plan should to develop and adopt to suit the ongoing change of demegraphic of homeless population.
 

### 2024-09-14 scripts/02-data_cleaning.R

@@ -11,11 +11,10 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(stringr)
 
 #### Clean data ####
 raw_data <- read_csv("inputs/data/raw_data.csv")
-
+c
 cleaned_data <-
   raw_data |>
   janitor::clean_names() 
@@ -42,7 +41,10 @@ data$Quarter <- paste0(year(data$date_mmm_yy), "Q", quarter(data$date_mmm_yy))
 # Group by Quarter and population_group, then calculate the average for all other columns
 quarterly_data <- data %>%
   group_by(Quarter, population_group) %>%
-  summarise(across(actively_homeless:Pre_Retirement_and_Retirement_Age, mean, na.rm = TRUE))
+  summarise(across(actively_homeless:Pre_Retirement_and_Retirement_Age, mean, 
+                   na.rm = TRUE))
 
+quarterly_data <- quarterly_data |> 
+  filter(Quarter != '2024Q3')
 #### Save data ####
 write_csv(quarterly_data, "outputs/data/analysis_data.csv")

### 2024-09-15 outputs/paper/paper.qmd
Second edition of intro and data, upadted citation.

@@
 title: "My title"
-subtitle: "My subtitle if needed"
 author: 
   - Junbo Li
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Junbo345/Homeless."
 date: today
@@
 #### Package setup ####
-
+library(here)
 library(tidyverse)
-library(here)
-library(tinytex)
-library(dplyr)
-library(tibble)
 library(opendatatoronto)
-library(janitor)
-library(ggplot2)
 library(knitr)
@@
 # Introduction
-According to @canadianobservatory2012definition, homeless is defined as "the situation of an individual, family or community without stable, safe, permanent, appropriate housing, or the immediate prospect, means and ability of acquiring it".
-It is estimated that about 23,500 Canadian people experienced every type of Homeless every year @gaetz2014state.
-On SEP 12, 2024, there are still more than 9500 people staying in shelters in the city of Toronto @toronto2024shelter.
-To reduce homelessness, Toronto city government published the supportive houseing development plan, which aims to build more affordable housings for the homeless people @toronto2024affordable.
-A recent two-year plan starting 2022 aiming to build 4,000 more affordable homes as well as supportive houseings by the end of 2024.
+Homelessness is defined as "the situation of an individual, family or community without stable, safe, permanent, appropriate housing, or the immediate prospect, means and ability of acquiring it" [@canadianobservatory2012definition].
+According to @gaetz2014state, approximately 23,500 Canadian people experience every type of homelessness every year.
+On SEP 12, 2024, there are still more than 9500 people staying in shelters in the city of Toronto [@toronto_shelter_census].
+To address this ongoing crisis, the Toronto city government launched the supportive housing development plan, which aims to build more affordable housing for homeless people [@toronto2024affordable].
+A recent two-year plan starting in 2022 aims to build 4,000 more affordable homes as well as supportive housing by the end of 2024 [@toronto2024affordable].
 
-In this paper, we studied the demographical shift of total homeless populations in the city of Toronto, and find that the underlyin demographics of homesless populations changed significantly.
-This would indicate that the tradition plan which mainly focus on buliding houses would no longer solve new homeless problem.
+While the city of Toronto has concentrated its efforts on building more affordable housing, our analysis in this paper has found that the underlying demographics of homeless populations changed significantly.
+This would indicate that the traditional housing-centric plan would no longer solve the newly evolved homeless problem.
 
-In the following sections, we will first introduce our sourse of data as well as potential cleaning procedures.
-Then, we will show our results of the demographic change.
-Finally, we would indicate the reason of our claim base on the analysis of the result and give some suggestions that would help better reduce homelessness in the city.
+In this paper, we delve into the demographic shift of Toronto's homeless population in greater detail by studying the monthly data from Toronto's shelter system from 2018 Jan to 2024 Jun [@toronto_shelter_data].
+We start by outlining the data source, the way we accessed the data, as well as procedures we used to clean and prepare the data in analysis in **Data** section.
+Following this, in **Result** section, we draw the bar graph demonstrating the trend in demographic changes and present our key findings.
+Finally, in **Discussion** section, we analyze the implications of the trends we found and offer recommendations for policy changes that would closely address the needs of different demographic groups.
 
@@
 
-The data used in here is downloaded from opendattoronto [@citeOpenDataToronto]. Specifically, we used the Shelter System Flow Data
-released by Toronto government [@toronto_shelter_data].The data loading, cleaning and analysis process is done by R [@citeR], togehter with packges include tidyverse [@tidy]...
-This data measures number of people requiring for shelter inside city of Toronto from Jan-2018 till Jun-2024 by each month (actively_homeless).
-Also, this dataset provides division of age group, refugee status, gender, as well as chronic status to provide further detailed study into the change of composition of homeless people over time.
-While there are several other dataset including daily shelter occupacancy in opendatatoronto, they do not have the subdivision information such as age or refugee status as this one.
+The data used here is downloaded from opendatatoronto [@citeOpenDataToronto].
+Specifically, we used the Shelter System Flow Data released by the Toronto government [@toronto_shelter_data].
+The data loading, cleaning, and analysis process is done by R [@citeR], together with packages including tidyverse [@tidy] here [@here], and knitr [@knitr].
+This data tracks the number of people requiring shelter inside the city of Toronto from Jan 2018 until Jun 2024 monthly (variable named **actively_homeless**).
+Also, this data set provides detailed demographic breakdowns such as age group, refugee status, gender, and chronic status allowing us to deploy analysis into the change of composition of homeless people over time.
+While there are several other data sets including daily shelter occupancy in opendatatoronto, the lack of detailed demographic information makes them unsuitable for our study.
 
@@
 
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
 # Filter the data for 'All Population' group
@@
 
-Second, due to increased global conflict in Eastern Europe and the Middle East, the number of refugees entering Toronto has increased @cbc23.On the other hand, the Toronto government has not been prepared to support such a huge increase of refugees as the portion of refugees in homeless people roared from less than 20% in the first three quarters of 2021 to nearly 60% at second quarter 2024.
+Second, due to increased global conflict in Eastern Europe and the Middle East, the number of refugees entering Toronto has increased [@cbc23].On the other hand, the Toronto government has not been prepared to support such a huge increase of refugees as the portion of refugees in homeless people roared from less than 20% in the first three quarters of 2021 to nearly 60% at second quarter 2024.
 We observe an emergency for the government to increase its speed of helping these newly entered refugees and give them a job and a secure place to live.

### 2024-09-15 outputs/paper/paper.qmd
Second edition of results

@@
 
+\newpage
+
 # Introduction
@@
 
-We dropeed the only missing row corresponding to july-2024.
-We selected our columns of interest, namely, gender: ('male', 'female', 'transgender_non_binary_or_two_spirit'), total number of homeless people every month ('actively_homeless'), as well as refugee and chronic status in polulation column.
-In addition, we grouped age group by whether they are suitable for work (age under 24 for early career stage, age25-44 for working age and age above 45 for pre-retirement and retirement age).
-Finally, we replace the month data by averaged quatered data for better analysis and visualization of a prelonged time period.
-The cleaned data could be seen in @tbl-cleandata.
+We dropped the only missing row corresponding to July 2024.
+We then selected our columns of interest, namely, the total number of homeless people every month **actively_homeless**, gender ratio: **male**, **female**, **transgender_non_binary_or_two_spirit**, as well as refugee and chronic status in column **population**.
+In addition, we categorize age groups by whether they are suitable for work: transform age under 24 for **early career stage**, age between 25 and 44 for **working age**, and age above 45 for **pre-retirement and retirement age**.
+Finally, we replace the monthly data with averaged quarterly data for better analysis and visualization of a prolonged period.
+A sample of cleaned data can be seen in @tbl-cleandata.
 
@@
 
-First, we plot the bar graph of over all trend of homeless people over the pass of the time.
-In @fig-overall, we observe an slightly increase of total number of homeless people from first quarter 2018 to second quarter of 2024.
-There is a decrease of homeless population in late 2020-early 2021, but increased quickly after that.
+First, we plot a bar graph to show the trend of the size of the entire homeless population of homeless people over the past time.
+In @fig-overall, we observe a slight increase in the total number of homeless people from the first quarter of 2018 to the second quarter of 2024, while in between there is a decrease of the homeless population in late 2020 to early 2021, followed by a rapid increase.
 
@@
 ## Gender trend
-Next, we investigate the trend of gender proportion.
-We plot the proportion of male, female, and Other genders against time in @fig-gender.
-Even thought there is some flatuations of the proportion of these three gender types, the overall trend is quite stable.
-The male proportion is around 60% while female proportion around 40% and others around 1%.
+Next, we investigate the trend of gender ratio.
+We plot the proportion of male, female, and Other gender types against time in @fig-gender.
+Even though there is some fluctuation in the proportion of these three gender types, the overall trend is stable.
+The male proportion stays around 60% while the female proportion stays around 40% and others around 1%.
 
 ## Chronic trend
-@fig-Chronic shows the trend of the proportion of chronic population among all of homeless people over the quarters of a year.
-Here, according to @chronic, chronic homeless refers to, "People who meet one of the two following criteria, as per the federal definition of chronic homelessness.
-The person has recorded a minimum of 180 overnight stay in the past year (365 days); or the person has recurrent overnight stays over the past three years with a cumulative duration of at least 546 nights." We can see form the graph that the proportion in chronic homeless increased a lot, from less than 35% in first quarter of 2018 to more than 65% in the second quarter of 2024.
+@fig-Chronic shows the trend of the proportion of chronic population among all homeless people over the quarters of each year.
+Here, according to [@chronic], chronic homelessness refers to "People who meet one of the two following criteria, as per the federal definition of chronic homelessness.
+The person has recorded a minimum of 180 overnight stays in the past year (365 days), or the person has recurrent overnight stays over the past three years with a cumulative duration of at least 546 nights." We can see from the graph that the proportion of chronic homeless increased a lot, from less than 35% in the first quarter of 2018 to more than 65% in the second quarter of 2024.
 ```{r}
@@
 ## Age trend
-@fig-WorkingAge shows the trend of the proportion of different working age statuses among the actively homeless population over the quarters of a year.
-Here, the working age statuses are categorized into three groups: Early Career Stage, Working Age, and Pre-Retirement and Retirement Age.Even thought the trend is not huge, We can see from the graph that the proportion of the Early Career Stage population  as well as pre-retire and retire have decreased (from 29% and 33% separately in first quarter of 2018 to 25% and 28% in second quarter in 2024).
-On the other hand, working age proportion increased to 47% in 2024Q2 from 37% in 2018Q1.
+@fig-WorkingAge displays the trend of the proportion of different working age statuses among the actively homeless population over the quarters of each year.
+Here, the working age statuses are categorized into three groups: Early Career Stage, Working Age, and Pre-Retirement and Retirement Age.
+Even though the trend is not dramatic, We can see from the graph that the proportion of the Early Career Stage population as well as pre-retire and retired has decreased (from 29% and 33% separately in the first quarter of 2018 to 25% and 28% in the second quarter in 2024).
+On the other hand, the working age proportion increased to 47% in 2024Q2 from 37% in 2018Q1.
+This indicates the growing population who are in the working age cannot afford their housing.
 ```{r}
@@
 ## Refugee trend
-@fig-refugee shows the trend of the proportion of refugees among the actively homeless population over the quarters of a year.
-We observe a signifiacant decrease of the trend accompanyed by even more signifcant increase.
+@fig-refugee shows the trend of the proportion of refugees among the actively homeless population over the quarters of each year.
+We observe a significant decrease in the proportion of individuals with refugee status from 35% in 2019 to less than 20% in 2021.
+However, starting in the year 2022, the proportion surged to nearly 60% in the second quarter of 2024.
 ```{r}

### 2024-09-18 outputs/paper/paper.qmd
second edition of discussion section

@@
 # Discussion
-Based on the analysis of the results section, we observe that, even though the overall population of homeless people did not incur significant variation in the past six years, its demographics changed a lot.
-Specifically, the proportion of refugees, chronic people as well as working-aged people.
-In the past, the City of Toronto has devoted its efforts to reducing the number of homeless people, especially through projects of building affordable houses [@torontosol].
-However, our study shows that with the change of demographics, the city government should also give its focus on other strategies for reducing homelessness.
+Based on the analysis of the results section, we observe that, even though the overall homeless population has remained relatively stable over the past six years, its demographics have significantly shifted.
+Specifically, the proportion of refugees, chronically homeless individuals as well as working-aged people among all homeless populations have increased.
+Historically, the City of Toronto has devoted its efforts to reducing the number of homeless people primarily through building more affordable houses [@torontosol].
+However, our study suggests that with these changes in demographics, the city government should expand its focus to address homelessness more effectively.
 
-First, as the result section points out, the proportion of chronically homeless people nearly doubled in the past six years.
-Despite its nature of long-lasting homeless status, studies have also shown that chronic homelessness is closely associated with chronic illness or mental problems @chronic.
-As pointed out by another study, the average income of people experiencing mental illness and other types of disorders is significantly lower than that of normal people @Sareen2011.
-Thus, it is very hard for those people who have already experienced mental and physical illness to escape homelessness.
-In addition, homeless people would face more severe health problems due to lack of adequate food, discrimination, barriers to health care, etc [@Sleet2021].
-A combination of these would become an obvious negative cycle: illness leads to a decrease in income, which leads to becoming homeless, which would further depend on illness.
-What the city government can do, on the other hand, is to increase the public health care system, especially focused on low-income and homeless people.
-In fact, merely building houses only increased the proportion of chronically homeless people and its aim is to reduce [@house_not].
+First, as the result section points out, the proportion of the chronically homeless population nearly doubled in the past six years.
+Despite its nature of long-lasting homeless status, studies have also shown that chronic homelessness is closely associated with chronic illness or other mental issues @chronic.
+Research further indicates that the average income of people experiencing mental illness and different types of disorders is significantly lower than normal people's income @Sareen2011.
+This makes it particularly difficult for people who have already experienced health problems to escape homelessness.
+In addition, homelessness exacerbates health problems due to lack of adequate food, discrimination, barriers to health care, etc.
+[@Sleet2021].
+A combination of these would become a harmful cycle: illness leads to a decrease in income, which leads to becoming homeless, which would further worsen individual illness.
 
-Second, due to increased global conflict in Eastern Europe and the Middle East, the number of refugees entering Toronto has increased [@cbc23].On the other hand, the Toronto government has not been prepared to support such a huge increase of refugees as the portion of refugees in homeless people roared from less than 20% in the first three quarters of 2021 to nearly 60% at second quarter 2024.
-We observe an emergency for the government to increase its speed of helping these newly entered refugees and give them a job and a secure place to live.
+To break the harmful cycle, the city government should give more attention to public health care systems, especially for low-income and homeless populations.
+In fact, merely building houses only increased the proportion of chronically homeless populations [@house_not].
+The city should, instead, shift to building supportive housing, where health services as well as social support are integrated with affordable housing.
+Mobile outreach teams and community-based mental health clinics can also ensure that homeless individuals receive the care they need.
+
+Second, due to increased global conflict in Eastern Europe and the Middle East, the number of refugees entering Toronto has increased [@cbc23].On the other hand, the city government has not been prepared to support such a huge increase of refugees as the portion of refugees homeless people roared from less than 20% in the first three quarters of 2021 to nearly 60% in the second quarter of 2024.
+We observed an emergency for the government to increase its speed in helping these newly entered refugees adapt to the new environment and give them a livable income job to support their housing.
 
 Finally, we observe an increase in the proportion of people of working age.
-This indicates the need for the government to create more jobs for the people who need them to make a living, like people who are experiencing mental health problems or physical disabilities.
+This indicates the need for the government to create more jobs for the low-income and homeless population, particularly for those with mental health issues or physical disabilities who rely on employment to sustain themselves.
 

### 2024-09-18 outputs/paper/paper.qmd
Second edition of weakness

@@
 
-Even though our data is the most uptodate and accurate date aviliable for homeless population on opendatatoronto, it might have potential bias towards estimation the ture homeless population in city of Toronto.
-One clear bias is that people would not stay in the shelter everyday and they would enter and exit the shelter frequently [@toronto_shelter_data].
-Correctly measure the ture popution demegraphic is a very diffcult on-going problem in Canada, but the shelter flow data would provide a proper estimate of our interest [@dionne2023review].
-To understand the probelms better and develop more strategies the next steps would be to research stratgies that could reach out and survey the population that is diffcult to contact such as unsheltered homeless population.
-Also, a new homeless reduction plan should to develop and adopt to suit the ongoing change of demegraphic of homeless population.
+Even though our data is the most up-to-date and accurate data available for the homeless population on Opendatatoronto, it may still have a bias toward the estimation of the true homeless population in the city of Toronto.
+For example, people would not stay in the shelter every day and they would enter and exit the shelter frequently [@toronto_shelter_data].
+Accurately measuring the true population demographic is a very difficult topic even in Canada, but the shelter flow data would provide a proper estimate of our research interest [@dionne2023review].
+To understand the problems better and develop more strategies, the next steps would be to research sampling methods that could reach out and survey the population that is difficult to contact such as the unsheltered homeless population.
+More incorporative sampling strategies would provide us better understanding of homeless demographics and develop further sophisticated plans to reduce homelessness.
 

### 2024-09-18 outputs/paper/paper.qmd
gramar chenge

@@
 # Discussion
-Based on the analysis of the results section, we observe that, even though the overall homeless population has remained relatively stable over the past six years, its demographics have significantly shifted.
+Based on the analysis of the results section, we observe that, even though the overall homeless population has remained relatively stable over the past six years, its demographics have shifted.
 Specifically, the proportion of refugees, chronically homeless individuals as well as working-aged people among all homeless populations have increased.
-Historically, the City of Toronto has devoted its efforts to reducing the number of homeless people primarily through building more affordable houses [@torontosol].
+Starting at 2020, the City of Toronto has devoted its efforts to reducing the number of homeless people primarily through building more affordable houses [@torontosol].
 However, our study suggests that with these changes in demographics, the city government should expand its focus to address homelessness more effectively.

### 2024-09-21 scripts/00-simulate_data.R
simulation with test case

@@ -1,19 +1,85 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates homeless population data
+# Author: Junbo Li
+# Date: SEP-20-2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr)
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# Load necessary libraries
+
+# Set the seed for reproducibility
+set.seed(123)
+
+# Create a date sequence from 2018-Jan to 2024-Jan (monthly data)
+dates <- seq(as.Date("2018-01-01"), as.Date("2024-01-01"), by = "month")
+
+# Simulate the total population as a normally distributed random variable
+mean_population <- 10000
+sd_population <- 1000
+total_population <- round(rnorm(length(dates), mean = mean_population, sd = sd_population))
+
+# Ensure total population is positive
+total_population[total_population < 0] <- 0
+
+# Simulate working status components
+# These three categories should sum up to the total population
+pre_working <- round(total_population * runif(length(dates), min = 0.2, max = 0.4))
+working <- round(total_population * runif(length(dates), min = 0.4, max = 0.6))
+retirement <- total_population - pre_working - working
+
+# Simulate chronic cases (chronic number is below total population)
+chronic <- round(runif(length(dates), min = 0.01, max = 0.99) * total_population)
+
+# Simulate refugee cases (refugee number is below total population)
+refugee <- round(runif(length(dates), min = 0.01, max = 0.99) * total_population)
+
+# Create a dataframe with all the data
+population_data <- data.frame(
+  Date = dates,
+  Total_Population = total_population,
+  Pre_Working = pre_working,
+  Working = working,
+  Retirement = retirement,
+  Chronic = chronic,
+  Refugee = refugee
+)
+
+# View the first few rows of the dataframe
+head(population_data)
+
+# Test 1: Ensure all entries are greater than 0
[…]

### 2024-09-23 paper/paper.qmd
changed repo structure

--- 2024-09-18 outputs/paper/paper.qmd
@@
 ---
-title: "My title"
+title: "Demographic Shifts in Toronto's Homeless Population: Evaluating and Reassessing City Solutions"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Reducing homeless population in city of Toronto is an ongoing task.
+While the reduce method by the government stays the same, the demographic of the population varys in the past six years.
+In our paper, we analysis the demographic change of homless population and show that the traditional solution is no longer effective.
+We will then present new solutions of better adapt the on-going shift of demographics."
 format: pdf

### 2024-09-23 scripts/02-data_cleaning.R
changed script for repo structure change

@@ -13,7 +13,7 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/raw_data.csv")
+raw_data <- read_csv("data/raw_data.csv")
 c
 cleaned_data <-
   raw_data |>
@@ -47,4 +47,4 @@ quarterly_data <- data %>%
 quarterly_data <- quarterly_data |> 
   filter(Quarter != '2024Q3')
 #### Save data ####
-write_csv(quarterly_data, "outputs/data/analysis_data.csv")
+write_csv(quarterly_data, "data/analysis_data.csv")

### 2024-09-24 paper/paper.qmd
abandon open a seperate pasge after toc
Changed repo structure
Deleted gender analysis due to on trend observed

@@
 # Read data
-raw_lead_data <- read_csv(here("inputs/data/raw_data.csv"))
-cleaned_data <- read_csv(here("outputs/data/analysis_data.csv"))
+raw_lead_data <- read_csv(here("data/raw_data/raw_data.csv"))
+cleaned_data <- read_csv(here("data/analysis_data/analysis_data.csv"))
 ```
-
-\newpage
 
@@
 We dropped the only missing row corresponding to July 2024.
-We then selected our columns of interest, namely, the total number of homeless people every month **actively_homeless**, gender ratio: **male**, **female**, **transgender_non_binary_or_two_spirit**, as well as refugee and chronic status in column **population**.
+We then selected our columns of interest, namely, the total number of homeless people every month **actively_homeless**, as well as refugee and chronic status in column **population**.
 In addition, we categorize age groups by whether they are suitable for work: transform age under 24 for **early career stage**, age between 25 and 44 for **working age**, and age above 45 for **pre-retirement and retirement age**.
@@
   kable(
-    col.names = c("Time (Quaters)", "Subdivision", "Total people", "Female", "Male", "Other Gender",
+    col.names = c("Time (Quaters)", "Subdivision", "Total people",
                   "Early Career Stage", "Working Age", "Pre-retire and Retire"),
@@
 ```
-
-```{r}
-#| label: fig-gender
-#| echo: false
-#| fig-cap: "Proportion of Active Homeless Population by Gender Over Quarters"
-
-# Filter the data for 'All Population' group
-all_population_data <- subset(cleaned_data, population_group == 'All Population')
-
-# Calculate the proportion of each gender
-all_population_data <- all_population_data %>%
-  mutate(total = gender_female + gender_male + gender_transgender_non_binary_or_two_spirit,
-         female_proportion = gender_female / total,
-         male_proportion = gender_male / total,
-         transgender_proportion = gender_transgender_non_binary_or_two_spirit / total)
-
-# Reshape the data for plotting
-proportion_data_long <- all_population_data %>%
-  select(Quarter, female_proportion, male_proportion, transgender_proportion) %>%
-  gather(key = "Gender", value = "Proportion", -Quarter)
-
-# Rename gender categories for better readability
-proportion_data_long$Gender <- recode(proportion_data_long$Gender,
-                                      female_proportion = "Female",
-                                      male_proportion = "Male",
-                                      transgender_proportion = "Other")
-
-# Create the stacked bar graph
-ggplot(proportion_data_long, aes(x = Quarter, y = Proportion, fill = Gender)) +
-  geom_bar(stat = "identity") +
-  labs(x = "Quarters of a year",
-       y = "Proportion") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  theme(legend.position = "bottom") +
-  scale_fill_manual(values = c("Female" = "pink", "Male" = "blue", "Other" = "purple"))
-
-```
-
-## Gender trend
-Next, we investigate the trend of gender ratio.
-We plot the proportion of male, female, and Other gender types against time in @fig-gender.
-Even though there is some fluctuation in the proportion of these three gender types, the overall trend is stable.
-The male proportion stays around 60% while the female proportion stays around 40% and others around 1%.
 

### 2024-09-24 scripts/01-download_data.R

@@ -30,6 +30,6 @@ data <- filter(datastore_resources, row_number()==1) %>% get_resource()
 
 
 #### Save data ####
-write_csv(data, "data/raw_data.csv") 
+write_csv(data, "data/raw_data/raw_data.csv") 
 
          

### 2024-09-24 scripts/02-data_cleaning.R

@@ -13,7 +13,7 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("data/raw_data.csv")
+raw_data <- read_csv("data/raw_data/raw_data.csv")
 c
 cleaned_data <-
   raw_data |>
@@ -47,4 +47,4 @@ quarterly_data <- data %>%
 quarterly_data <- quarterly_data |> 
   filter(Quarter != '2024Q3')
 #### Save data ####
-write_csv(quarterly_data, "data/analysis_data.csv")
+write_csv(quarterly_data, "data/analysis_data/analysis_data.csv")

### 2024-09-26 paper/paper.qmd
Updated abstract

@@
 date-format: long
-abstract: "Reducing homeless population in city of Toronto is an ongoing task.
-While the reduce method by the government stays the same, the demographic of the population varys in the past six years.
-In our paper, we analysis the demographic change of homless population and show that the traditional solution is no longer effective.
-We will then present new solutions of better adapt the on-going shift of demographics."
+abstract: "Reducing the homeless population in the City of Toronto is an ongoing task, and every year the city government struggles to address the different needs of homeless populations.
+In our paper, we analyze the demographic change of the homeless population in the past six years with data from Toronto's shelter system.
+We show that the proportion of chronic, refugees, and working-age all rose in the past six years.
+These findings indicate the city's current solution which focuses on building houses is no longer effective.
+We further propose new policies like improving public health systems to beat the latest crisis."
 format: pdf
@@
 
+# Give the summary table
 cleaned_data |>

### 2024-09-26 paper/paper.qmd
add variable description

@@
 Following this, in **Result** section, we draw the bar graph demonstrating the trend in demographic changes and present our key findings.
-Finally, in **Discussion** section, we analyze the implications of the trends we found and offer recommendations for policy changes that would closely address the needs of different demographic groups.
+Finally, in **Discussion** section, we analyze the implications of the trends we found and offer recommendations for policy changes that would address the needs of different demographic groups.
 
@@
 While there are several other data sets including daily shelter occupancy in opendatatoronto, the lack of detailed demographic information makes them unsuitable for our study.
+
+The Variables that is of our interest are:
+
+**1. date(mmm-yy)**: The time period the data was published (month/year, eg. Jan-18 means January of year 2018).
+
+**2. population_group**: Each row represent different population groups, here we focus on all population, chronic, and refugee.
+
+**3.
+actively_homeless**: This is the total number of people who have used the shelter system at least one time in the past 3 months.
+We will use this as an estimate of the total number of homeless population.
+
+**4. ageunder16**, **age16-24**, **age25-34**, **age35-44**, **age45-54**, **age55-64**, **age65over**: These are the number of homeless individuals categorized by their age.
 

### 2024-09-26 paper/paper.qmd
changed figure doc to be more clear

@@
 #| echo: false
-#| fig-cap: "Homeless Population Over Quarters"
+#| fig-cap: "Total homeless population in toronto, measured quarterly"
 
@@
 #| echo: false
-#| fig-cap: "Proportion of Chronic population to Active Population by Quarter"
+#| fig-cap: "Proportion of chronic homeless population to toal homeless population in Toronto, measured quarterly"
 
@@
   geom_bar(stat = "identity", fill = "blue") +
-  labs(x = "Quarters", y = "Proportion of Chronic to Active Population") +
+  labs(x = "Quarters of a year", y = "Proportion of Chronic to Total Homeless Population") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 #| echo: false
-#| fig-cap: "Proportion of working age population to Active Population by Quarter"
+#| fig-cap: "Proportion of working age population to to toal homeless population in Toronto, measured quarterly"
 
@@
 #| echo: false
-#| fig-cap: "Proportion of Chronic population to Active Population by Quarter"
+#| fig-cap: "Proportion of Chronic population to to to toal homeless population in Toronto, measured quarterly"
 

### 2024-09-26 scripts/00-simulate_data.R
updated simulation script

@@ -22,6 +22,12 @@ set.seed(123)
 # Create a date sequence from 2018-Jan to 2024-Jan (monthly data)
 dates <- seq(as.Date("2018-01-01"), as.Date("2024-01-01"), by = "month")
 
+# Repeat each date 3 times
+repeated_dates <- rep(dates, each = 3)
+
+# Format the dates in "MMM-YY" format
+formatted_dates <- format(repeated_dates, "%b-%y")
+
 # Simulate the total population as a normally distributed random variable
 mean_population <- 10000
 sd_population <- 1000
@@ -30,27 +36,30 @@ total_population <- round(rnorm(length(dates), mean = mean_population, sd = sd_p
 # Ensure total population is positive
 total_population[total_population < 0] <- 0
 
-# Simulate working status components
-# These three categories should sum up to the total population
-pre_working <- round(total_population * runif(length(dates), min = 0.2, max = 0.4))
-working <- round(total_population * runif(length(dates), min = 0.4, max = 0.6))
-retirement <- total_population - pre_working - working
-
 # Simulate chronic cases (chronic number is below total population)
 chronic <- round(runif(length(dates), min = 0.01, max = 0.99) * total_population)
 
 # Simulate refugee cases (refugee number is below total population)
 refugee <- round(runif(length(dates), min = 0.01, max = 0.99) * total_population)
 
+# Combine the three columns into one long column alternately
+df <- data.frame(total_population, chronic, refugee)
+combined_population <- c(t(df))
+
+# Simulate working status components
+# These three categories should sum up to the total population
+pre_working <- round(combined_population * runif(length(dates)*3, min = 0.2, max = 0.4))
+working <- round(combined_population * runif(length(dates)*3, min = 0.4, max = 0.6))
+retirement <- combined_population - pre_working - working
+
 # Create a dataframe with all the data
 population_data <- data.frame(
-  Date = dates,
-  Total_Population = total_population,
-  Pre_Working = pre_working,
-  Working = working,
-  Retirement = retirement,
-  Chronic = chronic,
-  Refugee = refugee
+  "Date" = formatted_dates,
+  "Population_group" = rep(c("All Population", "Chronic", "Refugees"), times = length(dates)),
+  "actively_homeless" = combined_population,
+  "Pre_Working" = pre_working,
+  "Working" = working,
+  "Retirement" = retirement
 )
 
 # View the first few rows of the dataframe
[…]

### 2024-09-26 paper/paper.qmd
Updated scripts

@@
 
-# Results {#sec-results}
+## Results
 
-## Overall trend
+### Overall trend
 
@@
 
-## Chronic trend
+### Chronic trend
 @fig-Chronic shows the trend of the proportion of chronic population among all homeless people over the quarters of each year.
@@
 ```
-## Age trend
+### Age trend
 @fig-WorkingAge displays the trend of the proportion of different working age statuses among the actively homeless population over the quarters of each year.
@@
 
-## Refugee trend
+### Refugee trend
 @fig-refugee shows the trend of the proportion of refugees among the actively homeless population over the quarters of each year.

### 2024-09-26 scripts/00-simulate_data.R

@@ -22,25 +22,28 @@ set.seed(123)
 # Create a date sequence from 2018-Jan to 2024-Jan (monthly data)
 dates <- seq(as.Date("2018-01-01"), as.Date("2024-01-01"), by = "month")
 
-# Repeat each date 3 times
-repeated_dates <- rep(dates, each = 3)
+# Extract the year and quarter from the dates
+quarters <- paste0(year(dates), "Q", quarter(dates))
 
-# Format the dates in "MMM-YY" format
-formatted_dates <- format(repeated_dates, "%b-%y")
+# Keep only unique quarters (quarterly data)
+quarterly_data <- quarters[!duplicated(quarters)]
+
+# Repeat each date 3 times
+repeated_dates <- rep(quarterly_data, each = 3)
 
 # Simulate the total population as a normally distributed random variable
 mean_population <- 10000
 sd_population <- 1000
-total_population <- round(rnorm(length(dates), mean = mean_population, sd = sd_population))
+total_population <- round(rnorm(length(quarterly_data), mean = mean_population, sd = sd_population))
 
 # Ensure total population is positive
 total_population[total_population < 0] <- 0
 
 # Simulate chronic cases (chronic number is below total population)
-chronic <- round(runif(length(dates), min = 0.01, max = 0.99) * total_population)
+chronic <- round(runif(length(quarterly_data), min = 0.01, max = 0.99) * total_population)
 
 # Simulate refugee cases (refugee number is below total population)
-refugee <- round(runif(length(dates), min = 0.01, max = 0.99) * total_population)
+refugee <- round(runif(length(quarterly_data), min = 0.01, max = 0.99) * total_population)
 
 # Combine the three columns into one long column alternately
 df <- data.frame(total_population, chronic, refugee)
@@ -48,41 +51,57 @@ combined_population <- c(t(df))
 
 # Simulate working status components
 # These three categories should sum up to the total population
-pre_working <- round(combined_population * runif(length(dates)*3, min = 0.2, max = 0.4))
-working <- round(combined_population * runif(length(dates)*3, min = 0.4, max = 0.6))
+pre_working <- round(combined_population * runif(length(quarterly_data)*3, min = 0.2, max = 0.4))
+working <- round(combined_population * runif(length(quarterly_data)*3, min = 0.4, max = 0.6))
 retirement <- combined_population - pre_working - working
 
 # Create a dataframe with all the data
 population_data <- data.frame(
-  "Date" = formatted_dates,
[…]

### 2024-09-26 scripts/02-data_cleaning.R

@@ -24,8 +24,7 @@ cleaned_data1 <- cleaned_data |>
     Early_Career_Stage = ageunder16 + age16_24) |>
   mutate(Working_Age = age25_34 + age35_44) |>
   mutate(Pre_Retirement_and_Retirement_Age = age45_54 + age55_64 + age65over) |>
-  select(date_mmm_yy, population_group, actively_homeless, 
-         gender_female, gender_male, gender_transgender_non_binary_or_two_spirit,
+  select(date_mmm_yy, population_group, actively_homeless,
          Early_Career_Stage, Working_Age, Pre_Retirement_and_Retirement_Age) |>
   filter(population_group == "All Population" | population_group == "Chronic" |
            population_group == "Refugees") |>

### 2024-09-26 scripts/03-test_data.R

@@ -10,6 +10,65 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+data <- read.csv("data/analysis_data/analysis_data.csv")
+
+#Test 1: Ensure all cells are non-empty
+# Function to check if any cell in the data frame is empty (NA or "")
+check_empty_cells <- function(df) {
+  # Check for NA values
+  has_na <- any(is.na(df))
+  
+  # Check for empty strings in character columns
+  has_empty_strings <- any(sapply(df, function(x) {
+    if (is.character(x)) {
+      any(x == "")
+    } else {
+      FALSE
+    }
+  }))
+  
+  # Return TRUE if either NA or empty strings are found
+  return(has_na || has_empty_strings)
+}
+
+# Apply the function
+if (check_empty_cells(data)) {
+  print("The data frame contains empty cells (NA or empty strings).")
+} else {
+  print("The data frame has no empty cells.")
+}
+
+
+
+# Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
+test_sum_equals_total <- all(data$Early_Career_Stage + 
+                               data$Working_Age + 
+                               data$Pre_Retirement_and_Retirement_Age == data$actively_homeless)
+
+if(test_sum_equals_total) {
+  print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
+} else {
+  print("Test 2 Failed: The sum of the working categories does not equal the total population.")
+}
+
+
+#Test 3: Ensure all numerical cells are greater than 0 
+# Function to check if all numerical columns have values greater than 0
+test_numerical_gt_zero <- function(df) {
+  # Apply the condition only to numeric columns
+  numeric_columns <- sapply(df, is.numeric)
+  all(sapply(df[, numeric_columns], function(x) all(x > 0)))
+}
+
+# Apply the test
+result <- test_numerical_gt_zero(data)
+
+# Display the result
+if (result) {
+  print("All numerical inputs are greater than 0")
+} else {
+  print("Some numerical inputs are less than or equal to 0")
+}

### 2024-09-26 scripts/02-data_cleaning.R
changed heading of scripts

@@ -6,7 +6,7 @@
 # Date: 13 SEP 2024
 # Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: raw data downloaded
 # Any other information needed? None
 
 #### Workspace setup ####
@@ -47,3 +47,28 @@ quarterly_data <- quarterly_data |>
   filter(Quarter != '2024Q3')
 #### Save data ####
 write_csv(quarterly_data, "data/analysis_data/analysis_data.csv")
+
+# Test 1: Ensure all entries are greater than 0
+test_greater_than_zero <- all(quarterly_data$Total_Population > 0) &
+  all(quarterly_data$Pre_Working >= 0) &
+  all(quarterly_data$Working >= 0) &
+  all(quarterly_data$Retirement >= 0) &
+  all(quarterly_data$Chronic >= 0) &
+  all(quarterly_data$Refugee >= 0)
+
+if(test_greater_than_zero) {
+  print("Test 1 Passed: All entries are greater than or equal to 0.")
+} else {
+  print("Test 1 Failed: Some entries are less than 0.")
+}
+
+# Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
+test_sum_equals_total <- all(quarterly_data$Pre_Retirement_and_Retirement_Age + 
+                               quarterly_data$Working_Age + 
+                               quarterly_data$Early_Career_Stage == quarterly_data$actively_homeless)
+
+if(test_sum_equals_total) {
+  print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
+} else {
+  print("Test 2 Failed: The sum of the working categories does not equal the total population.")
+}

### 2024-09-26 scripts/03-test_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the cleaned date is neat and correct
+# Author: Junbo Li
+# Date: 13 SEP 2024
+# Contact: junb.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Cleaned analysis data
+# Any other information needed? None
 
 
 #### Workspace setup ####

### 2024-09-26 scripts/03-test_data.R
heading change

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the cleaned date is neat and correct
 # Author: Junbo Li
-# Date: 13 SEP 2024
+# Date: 26 SEP 2024
 # Contact: junb.li@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Cleaned analysis data

### 2024-09-26 paper/paper.qmd
add measurement

@@
 
-The data used here is downloaded from opendatatoronto [@citeOpenDataToronto].
-Specifically, we used the Shelter System Flow Data released by the Toronto government [@toronto_shelter_data].
-The data loading, cleaning, and analysis process is done by R [@citeR], together with packages including tidyverse [@tidy] here [@here], and knitr [@knitr].
-This data tracks the number of people requiring shelter inside the city of Toronto from Jan 2018 until Jun 2024 monthly (variable named **actively_homeless**).
-Also, this data set provides detailed demographic breakdowns such as age group, refugee status, gender, and chronic status allowing us to deploy analysis into the change of composition of homeless people over time.
-While there are several other data sets including daily shelter occupancy in opendatatoronto, the lack of detailed demographic information makes them unsuitable for our study.
+The data for this analysis was sourced from opendatatoronto [@citeOpenDataToronto], specifically utilizing the Shelter System Flow Data released by the Toronto government [@toronto_shelter_data].
+The City of Toronto employs the Shelter Management Information System (SMIS) to track individuals accessing overnight homelessness services, including shelters, respites, and hotel/motel programs.
+This dataset captures the number of unique individuals who have used the shelter system at least once in the past three months and are still experiencing homelessness, meaning they have not been discharged to permanent housing.
+
+Updated monthly, the dataset is published on the 15th of each month, with revisions made to the previous month's data to enhance discharge accuracy.
+However, it currently excludes individuals sleeping outdoors or those utilizing services not funded by the City, omitting an estimated 18% of the homeless population.
+The dataset provides detailed demographic information such as age group, refugee status, gender, and chronic status, enabling us to analyze changes in the homeless population's composition over time.
+
+While other datasets, such as daily shelter occupancy, are available, they lack the demographic depth necessary for this study.
+Data loading, cleaning, and analysis were carried out in R [@citeR], utilizing packages like tidyverse [@tidy], here [@here], and knitr [@knitr].
+This comprehensive approach ensures that our analysis accurately reflects the realities faced by individuals experiencing homelessness in Toronto.
 

### 2024-09-26 paper/paper.qmd
summary statistics added

@@
 Finally, we replace the monthly data with averaged quarterly data for better analysis and visualization of a prolonged period.
-A sample of cleaned data can be seen in @tbl-cleandata.
+A sample of cleaned data can be seen in @tbl-cleandata and summary statistics in @tbl-summary
 
@@
     digits = 0)
+```
+
+```{r}
+#| label: tbl-summary
+#| echo: false
+#| tbl-cap: "summary statistics of cleaned data"
+
+# Group by Subdivision and calculate summary statistics
+summary_statistics <- cleaned_data %>%
+  group_by(population_group) %>%
+  summarize(
+    Mean_Total_people = mean(actively_homeless, na.rm = TRUE),
+    Mean_Early_Career_Stage = mean(Early_Career_Stage, na.rm = TRUE),
+    Mean_Working_Age = mean(Working_Age, na.rm = TRUE),
+    Mean_Pre_Retire_and_Retire = mean(Pre_Retirement_and_Retirement_Age, na.rm = TRUE)
+  )
+
+# Display the summary statistics using kable
+summary_statistics |>
+  kable(
+    col.names = c("Subdivision", "Mean Total People", 
+                  "Mean Early Career Stage", 
+                  "Mean Working Age", 
+                  "Mean Pre-retire and Retire"
+  ),
+    digits = 2
+  )
 

### 2024-09-27 paper/paper.qmd
changed pic style, changed age measurement from 25-44 to 25-54

@@
 We dropped the only missing row corresponding to July 2024.
-We then selected our columns of interest, namely, the total number of homeless people every month **actively_homeless**, as well as refugee and chronic status in column **population**.
-In addition, we categorize age groups by whether they are suitable for work: transform age under 24 for **early career stage**, age between 25 and 44 for **working age**, and age above 45 for **pre-retirement and retirement age**.
-Finally, we replace the monthly data with averaged quarterly data for better analysis and visualization of a prolonged period.
-A sample of cleaned data can be seen in @tbl-cleandata and summary statistics in @tbl-summary
+We then selected our columns of interest, namely, the total number of homeless people every month (**actively_homeless**), as well as **refugee** and **chronic status** in the column population.
+In addition, we categorized age groups based on their suitability for work: transforming ages under 24 for the **early career stage**, ages between 25 and 54 for **working age**, and ages above 55 for **pre-retirement and retirement age**.
+The classification of ages 25-54 as working age is aligned with the Canadian government's survey, which defines individuals aged 25-54 as core-aged workers @statcan2024.
+Finally, we replaced the monthly data with averaged quarterly data for better analysis and visualization over a prolonged period.
+A sample of cleaned data can be seen in @tbl-cleandata, and summary statistics in @tbl-summary.
 
@@
 #| echo: false
-#| fig-cap: "Total homeless population in toronto, measured quarterly"
+#| fig-cap: ""
 
@@
   labs(x = "Quarters of a year",
-       y = "Total number of homeless people") +
+       y = "Total number of homeless people", 
+       title = "Total homeless population in toronto, measured quarterly") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 #| echo: false
-#| fig-cap: "Proportion of chronic homeless population to toal homeless population in Toronto, measured quarterly"
+#| fig-cap: ""
 
@@
   geom_bar(stat = "identity", fill = "blue") +
-  labs(x = "Quarters of a year", y = "Proportion of Chronic to Total Homeless Population") +
+  labs(x = "Quarters of a year", y = "Proportion",
+       title = "Proportion of chronic homeless population to toal homeless\npopulation in Toronto, measured quarterly") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 Here, the working age statuses are categorized into three groups: Early Career Stage, Working Age, and Pre-Retirement and Retirement Age.
-Even though the trend is not dramatic, We can see from the graph that the proportion of the Early Career Stage population as well as pre-retire and retired has decreased (from 29% and 33% separately in the first quarter of 2018 to 25% and 28% in the second quarter in 2024).
-On the other hand, the working age proportion increased to 47% in 2024Q2 from 37% in 2018Q1.
+Even though the trend is not dramatic, We can see from the graph that the proportion of the Early Career Stage population as well as pre-retire and retired has decreased (from 29% and 17% separately in the first quarter of 2018 to 24.5% and 14.5% in the second quarter in 2024).
+On the other hand, the working age proportion increased to 61% in 2024Q2 from 54% in 2018Q1.
 This indicates the growing population who are in the working age cannot afford their housing.
@@
 #| echo: false
-#| fig-cap: "Proportion of working age population to to toal homeless population in Toronto, measured quarterly"
+#| fig-cap: ""
 
@@
   labs(x = "Quarters of a year",
-       y = "Proportion") +
+       y = "Proportion",
+       title = "Proportion of working age population to to toal homeless\npopulation in Toronto, measured quarterly") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom",
+        legend.text = element_text(size = 6)) +
   scale_fill_manual(values = c("Early Career Stage" = "blue", 
@@
 #| echo: false
-#| fig-cap: "Proportion of Chronic population to to to toal homeless population in Toronto, measured quarterly"
+#| fig-cap: ""
 
@@
   geom_bar(stat = "identity", fill = "blue") +
-  labs(x = "Quarters", y = "Proportion of Refugee to Active Population") +
+  labs(x = "Quarters", y = "Proportion",
+       title = "Proportion of Chronic population to to to toal homeless\npopulation in Toronto, measured quarterly") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2024-09-27 scripts/02-data_cleaning.R

@@ -22,8 +22,8 @@ cleaned_data <-
 cleaned_data1 <- cleaned_data |> 
   mutate(
     Early_Career_Stage = ageunder16 + age16_24) |>
-  mutate(Working_Age = age25_34 + age35_44) |>
-  mutate(Pre_Retirement_and_Retirement_Age = age45_54 + age55_64 + age65over) |>
+  mutate(Working_Age = age25_34 + age35_44 + age45_54) |>
+  mutate(Pre_Retirement_and_Retirement_Age = age55_64 + age65over) |>
   select(date_mmm_yy, population_group, actively_homeless,
          Early_Career_Stage, Working_Age, Pre_Retirement_and_Retirement_Age) |>
   filter(population_group == "All Population" | population_group == "Chronic" |
@@ -47,28 +47,3 @@ quarterly_data <- quarterly_data |>
   filter(Quarter != '2024Q3')
 #### Save data ####
 write_csv(quarterly_data, "data/analysis_data/analysis_data.csv")
-
-# Test 1: Ensure all entries are greater than 0
-test_greater_than_zero <- all(quarterly_data$Total_Population > 0) &
-  all(quarterly_data$Pre_Working >= 0) &
-  all(quarterly_data$Working >= 0) &
-  all(quarterly_data$Retirement >= 0) &
-  all(quarterly_data$Chronic >= 0) &
-  all(quarterly_data$Refugee >= 0)
-
-if(test_greater_than_zero) {
-  print("Test 1 Passed: All entries are greater than or equal to 0.")
-} else {
-  print("Test 1 Failed: Some entries are less than 0.")
-}
-
-# Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
-test_sum_equals_total <- all(quarterly_data$Pre_Retirement_and_Retirement_Age + 
-                               quarterly_data$Working_Age + 
-                               quarterly_data$Early_Career_Stage == quarterly_data$actively_homeless)
-
-if(test_sum_equals_total) {
-  print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
-} else {
-  print("Test 2 Failed: The sum of the working categories does not equal the total population.")
-}

### 2024-09-27 scripts/03-test_data.R

@@ -44,9 +44,9 @@ if (check_empty_cells(data)) {
 
 
 # Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
-test_sum_equals_total <- all(data$Early_Career_Stage + 
+test_sum_equals_total <- all(round(data$Early_Career_Stage + 
                                data$Working_Age + 
-                               data$Pre_Retirement_and_Retirement_Age == data$actively_homeless)
+                               data$Pre_Retirement_and_Retirement_Age) == round(data$actively_homeless))
 
 if(test_sum_equals_total) {
   print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")

### 2024-09-27 paper/paper.qmd
Updated code style using styler

@@
 In addition, we categorized age groups based on their suitability for work: transforming ages under 24 for the **early career stage**, ages between 25 and 54 for **working age**, and ages above 55 for **pre-retirement and retirement age**.
-The classification of ages 25-54 as working age is aligned with the Canadian government's survey, which defines individuals aged 25-54 as core-aged workers @statcan2024.
+The classification of ages 25-54 as working age is aligned with the Canadian government's survey, which defines individuals aged 25-54 as core-aged workers [@statcan2024].
 Finally, we replaced the monthly data with averaged quarterly data for better analysis and visualization over a prolonged period.
@@
   kable(
-    col.names = c("Time (Quaters)", "Subdivision", "Total people",
-                  "Early Career Stage", "Working Age", "Pre-retire and Retire"),
-    digits = 0)
+    col.names = c(
+      "Time (Quaters)", "Subdivision", "Total people",
+      "Early Career Stage", "Working Age", "Pre-retire and Retire"
+    ),
+    digits = 0
+  )
 ```
@@
   kable(
-    col.names = c("Subdivision", "Mean Total People", 
+    col.names = c(
+      "Subdivision", "Mean Total People",
       "Mean Early Career Stage",
@@
   )
-
 ```
@@
 # Filter the data for 'All Population' group
-all_population_data <- subset(cleaned_data, population_group == 'All Population')
+all_population_data <- subset(cleaned_data, population_group == "All Population")
 
@@
   geom_bar(stat = "identity", fill = "skyblue") +
-  labs(x = "Quarters of a year",
+  labs(
+    x = "Quarters of a year",
     y = "Total number of homeless people",
-       title = "Total homeless population in toronto, measured quarterly") +
+    title = "Total homeless population in toronto, measured quarterly"
+  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
 ```
@@
   geom_bar(stat = "identity", fill = "blue") +
-  labs(x = "Quarters of a year", y = "Proportion",
-       title = "Proportion of chronic homeless population to toal homeless\npopulation in Toronto, measured quarterly") +
+  labs(
+    x = "Quarters of a year", y = "Proportion",
+    title = "Proportion of chronic homeless population to toal homeless\npopulation in Toronto, measured quarterly"
+  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 all_population_data <- all_population_data %>%
-  mutate(total = Early_Career_Stage + Working_Age + Pre_Retirement_and_Retirement_Age,
+  mutate(
+    total = Early_Career_Stage + Working_Age + Pre_Retirement_and_Retirement_Age,
     early_career_proportion = Early_Career_Stage / total,
     working_age_proportion = Working_Age / total,
-         pre_retirement_proportion = Pre_Retirement_and_Retirement_Age / total)
+    pre_retirement_proportion = Pre_Retirement_and_Retirement_Age / total
+  )
 
@@
   working_age_proportion = "Working Age",
-                                          pre_retirement_proportion = "Pre-Retirement and Retirement Age")
+  pre_retirement_proportion = "Pre-Retirement and Retirement Age"
+)
 
@@
   geom_bar(stat = "identity") +
-  labs(x = "Quarters of a year",
+  labs(
+    x = "Quarters of a year",
     y = "Proportion",
-       title = "Proportion of working age population to to toal homeless\npopulation in Toronto, measured quarterly") +
+    title = "Proportion of working age population to to toal homeless\npopulation in Toronto, measured quarterly"
+  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  theme(legend.position = "bottom",
-        legend.text = element_text(size = 6)) +
-  scale_fill_manual(values = c("Early Career Stage" = "blue", 
+  theme(
+    legend.position = "bottom",
+    legend.text = element_text(size = 6)
+  ) +
+  scale_fill_manual(values = c(
+    "Early Career Stage" = "blue",
     "Working Age" = "pink",
-                               "Pre-Retirement and Retirement Age" = "green"))
+    "Pre-Retirement and Retirement Age" = "green"
+  ))
 ```
@@
   geom_bar(stat = "identity", fill = "blue") +
-  labs(x = "Quarters", y = "Proportion",
-       title = "Proportion of Chronic population to to to toal homeless\npopulation in Toronto, measured quarterly") +
+  labs(
+    x = "Quarters", y = "Proportion",
+    title = "Proportion of Chronic population to to to toal homeless\npopulation in Toronto, measured quarterly"
+  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2024-09-27 scripts/00-simulate_data.R

@@ -51,8 +51,8 @@ combined_population <- c(t(df))
 
 # Simulate working status components
 # These three categories should sum up to the total population
-pre_working <- round(combined_population * runif(length(quarterly_data)*3, min = 0.2, max = 0.4))
-working <- round(combined_population * runif(length(quarterly_data)*3, min = 0.4, max = 0.6))
+pre_working <- round(combined_population * runif(length(quarterly_data) * 3, min = 0.2, max = 0.4))
+working <- round(combined_population * runif(length(quarterly_data) * 3, min = 0.4, max = 0.6))
 retirement <- combined_population - pre_working - working
 
 # Create a dataframe with all the data
@@ -65,17 +65,17 @@ population_data <- data.frame(
   "Retirement" = retirement
 )
 
-write_csv(population_data, "data/raw_data/simulated_data.csv") 
+write_csv(population_data, "data/raw_data/simulated_data.csv")
 
 
-#Tests
+# Tests
 
-#Test 1: Ensure all cells are non-empty
+# Test 1: Ensure all cells are non-empty
 # Function to check if any cell in the data frame is empty (NA or "")
 check_empty_cells <- function(df) {
   # Check for NA values
   has_na <- any(is.na(df))
-  
+
   # Check for empty strings in character columns
   has_empty_strings <- any(sapply(df, function(x) {
     if (is.character(x)) {
@@ -84,7 +84,7 @@ check_empty_cells <- function(df) {
       FALSE
     }
   }))
-  
+
   # Return TRUE if either NA or empty strings are found
   return(has_na || has_empty_strings)
 }
@@ -99,18 +99,18 @@ if (check_empty_cells(population_data)) {
 
 
 # Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
-test_sum_equals_total <- all(population_data$Pre_Working + 
-                               population_data$Working + 
-                               population_data$Retirement == population_data$actively_homeless)
+test_sum_equals_total <- all(population_data$Pre_Working +
+  population_data$Working +
+  population_data$Retirement == population_data$actively_homeless)
 
-if(test_sum_equals_total) {
+if (test_sum_equals_total) {
   print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
 } else {
   print("Test 2 Failed: The sum of the working categories does not equal the total population.")
 }
 
 
-#Test 3: Ensure all numerical cells are greater than 0 
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -22,14 +22,12 @@ package
 resources <- list_package_resources("ac77f532-f18b-427c-905c-4ae87ce69c93")
 
 # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+datastore_resources <- filter(resources, tolower(format) %in% c("csv", "geojson"))
 
 # load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data <- filter(datastore_resources, row_number() == 1) %>% get_resource()
 
 
 
 #### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
-
-         
+write_csv(data, "data/raw_data/raw_data.csv")

### 2024-09-27 scripts/02-data_cleaning.R

@@ -17,22 +17,25 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 c
 cleaned_data <-
   raw_data |>
-  janitor::clean_names() 
+  janitor::clean_names()
 
-cleaned_data1 <- cleaned_data |> 
+cleaned_data1 <- cleaned_data |>
   mutate(
-    Early_Career_Stage = ageunder16 + age16_24) |>
+    Early_Career_Stage = ageunder16 + age16_24
+  ) |>
   mutate(Working_Age = age25_34 + age35_44 + age45_54) |>
   mutate(Pre_Retirement_and_Retirement_Age = age55_64 + age65over) |>
-  select(date_mmm_yy, population_group, actively_homeless,
-         Early_Career_Stage, Working_Age, Pre_Retirement_and_Retirement_Age) |>
+  select(
+    date_mmm_yy, population_group, actively_homeless,
+    Early_Career_Stage, Working_Age, Pre_Retirement_and_Retirement_Age
+  ) |>
   filter(population_group == "All Population" | population_group == "Chronic" |
-           population_group == "Refugees") |>
+    population_group == "Refugees") |>
   tidyr::drop_na()
 
 data <- cleaned_data1
 
-data$date_mmm_yy <- as.Date(paste0("01-", data$date_mmm_yy), format="%d-%b-%y")
+data$date_mmm_yy <- as.Date(paste0("01-", data$date_mmm_yy), format = "%d-%b-%y")
 
 # Create a new column for Quarters
 data$Quarter <- paste0(year(data$date_mmm_yy), "Q", quarter(data$date_mmm_yy))
@@ -40,10 +43,11 @@ data$Quarter <- paste0(year(data$date_mmm_yy), "Q", quarter(data$date_mmm_yy))
 # Group by Quarter and population_group, then calculate the average for all other columns
 quarterly_data <- data %>%
   group_by(Quarter, population_group) %>%
-  summarise(across(actively_homeless:Pre_Retirement_and_Retirement_Age, mean, 
-                   na.rm = TRUE))
+  summarise(across(actively_homeless:Pre_Retirement_and_Retirement_Age, mean,
+    na.rm = TRUE
+  ))
 
-quarterly_data <- quarterly_data |> 
-  filter(Quarter != '2024Q3')
+quarterly_data <- quarterly_data |>
+  filter(Quarter != "2024Q3")
 #### Save data ####
 write_csv(quarterly_data, "data/analysis_data/analysis_data.csv")

### 2024-09-27 scripts/03-test_data.R

@@ -15,12 +15,12 @@ library(tidyverse)
 #### Test data ####
 data <- read.csv("data/analysis_data/analysis_data.csv")
 
-#Test 1: Ensure all cells are non-empty
+# Test 1: Ensure all cells are non-empty
 # Function to check if any cell in the data frame is empty (NA or "")
 check_empty_cells <- function(df) {
   # Check for NA values
   has_na <- any(is.na(df))
-  
+
   # Check for empty strings in character columns
   has_empty_strings <- any(sapply(df, function(x) {
     if (is.character(x)) {
@@ -29,7 +29,7 @@ check_empty_cells <- function(df) {
       FALSE
     }
   }))
-  
+
   # Return TRUE if either NA or empty strings are found
   return(has_na || has_empty_strings)
 }
@@ -44,18 +44,18 @@ if (check_empty_cells(data)) {
 
 
 # Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
-test_sum_equals_total <- all(round(data$Early_Career_Stage + 
-                               data$Working_Age + 
-                               data$Pre_Retirement_and_Retirement_Age) == round(data$actively_homeless))
+test_sum_equals_total <- all(round(data$Early_Career_Stage +
+  data$Working_Age +
+  data$Pre_Retirement_and_Retirement_Age) == round(data$actively_homeless))
 
-if(test_sum_equals_total) {
+if (test_sum_equals_total) {
   print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
 } else {
   print("Test 2 Failed: The sum of the working categories does not equal the total population.")
 }
 
 
-#Test 3: Ensure all numerical cells are greater than 0 
+# Test 3: Ensure all numerical cells are greater than 0
 # Function to check if all numerical columns have values greater than 0
 test_numerical_gt_zero <- function(df) {
   # Apply the condition only to numeric columns

### 2024-09-27 paper/paper.qmd
cite styler

@@
 While other datasets, such as daily shelter occupancy, are available, they lack the demographic depth necessary for this study.
-Data loading, cleaning, and analysis were carried out in R [@citeR], utilizing packages like tidyverse [@tidy], here [@here], and knitr [@knitr].
+Data loading, cleaning, and analysis were carried out in **R** [@citeR], with packages **tidyverse** [@tidy], **here** [@here], **styler** [@style], and **knitr** [@knitr].
 This comprehensive approach ensures that our analysis accurately reflects the realities faced by individuals experiencing homelessness in Toronto.

### 2024-09-27 paper/paper.qmd
small changes for body
3rd edition of paper, with chat history updated

@@
 date-format: long
-abstract: "Reducing the homeless population in the City of Toronto is an ongoing task, and every year the city government struggles to address the different needs of homeless populations.
-In our paper, we analyze the demographic change of the homeless population in the past six years with data from Toronto's shelter system.
-We show that the proportion of chronic, refugees, and working-age all rose in the past six years.
-These findings indicate the city's current solution which focuses on building houses is no longer effective.
-We further propose new policies like improving public health systems to beat the latest crisis."
+abstract: "Homelessness in the City of Toronto has been a growing issue, with the city government facing difficulties in meeting the needs of different groups within the homeless population.
+In this paper, we analyze changes in the demographics of Toronto’s homeless population over the past six years, using data from the city's shelter system.
+We found that the proportion of individuals experiencing chronic homelessness, refugees, and working-age adults has increased.
+These findings suggest that the city's current focus on building housing is not enough, and we propose improvements in public health services to better address the evolving crisis."
 format: pdf
@@
 # Introduction
-Homelessness is defined as "the situation of an individual, family or community without stable, safe, permanent, appropriate housing, or the immediate prospect, means and ability of acquiring it" [@canadianobservatory2012definition].
-According to @gaetz2014state, approximately 23,500 Canadian people experience every type of homelessness every year.
-On SEP 12, 2024, there are still more than 9500 people staying in shelters in the city of Toronto [@toronto_shelter_census].
-To address this ongoing crisis, the Toronto city government launched the supportive housing development plan, which aims to build more affordable housing for homeless people [@toronto2024affordable].
-A recent two-year plan starting in 2022 aims to build 4,000 more affordable homes as well as supportive housing by the end of 2024 [@toronto2024affordable].
+Homelessness is defined as "the situation of an individual, family, or community without stable, safe, permanent, appropriate housing, or the immediate prospect, means, and ability of acquiring it" [@canadianobservatory2012definition].
+In Canada, around 23,500 people experience various forms of homelessness every year [@gaetz2014state], with more than 9,500 individuals currently living in shelters in Toronto as of September 12, 2024 [@toronto_shelter_census].
+To address this crisis, the Toronto city government has implemented several initiatives, including a supportive housing development plan, which aims to build more affordable housing for homeless populations [@toronto2024affordable].
+A key part of this plan is the 2022-2024 two-year effort to construct 4,000 new affordable homes [@toronto2024affordable].
 
-While the city of Toronto has concentrated its efforts on building more affordable housing, our analysis in this paper has found that the underlying demographics of homeless populations changed significantly.
-This would indicate that the traditional housing-centric plan would no longer solve the newly evolved homeless problem.
+While the city has made progress in expanding housing, we find that a housing-centric approach may not be sufficient to address the needs of the evolving homeless population.
+Our analysis reveals significant demographic changes in Toronto’s homeless population over the past six years, including increases in the proportion of chronically homeless individuals, refugees, and working-age adults.
+These demographic shifts suggest that the traditional emphasis on housing development alone is no longer be an adequate solution.
 
-In this paper, we delve into the demographic shift of Toronto's homeless population in greater detail by studying the monthly data from Toronto's shelter system from 2018 Jan to 2024 Jun [@toronto_shelter_data].
-We start by outlining the data source, the way we accessed the data, as well as procedures we used to clean and prepare the data in analysis in **Data** section.
-Following this, in **Result** section, we draw the bar graph demonstrating the trend in demographic changes and present our key findings.
-Finally, in **Discussion** section, we analyze the implications of the trends we found and offer recommendations for policy changes that would address the needs of different demographic groups.
+In this paper, we explore these demographic changes using monthly data from Toronto's Shelter Management Information System (SMIS) from January 2018 to June 2024 [@toronto_shelter_data].
+We detail the data collection process, cleaning methods, and key findings in the Data section.
+In the final section, Discussion, we assess the implications of our results for policy and propose recommendations for more comprehensive strategies that extend beyond housing solutions to address the diverse needs of Toronto's homeless population.
 
@@
 Finally, we replaced the monthly data with averaged quarterly data for better analysis and visualization over a prolonged period.
-A sample of cleaned data can be seen in @tbl-cleandata, and summary statistics in @tbl-summary.
+A sample of cleaned data can be seen in @tbl-cleandata, and the key features for each variable is in @tbl-summary.
 
@@
 
-## Results
+## Demographic trend with respect to time
 

### 2024-09-27 paper/paper.qmd
3rd edition of limitations

@@
 
-Even though our data is the most up-to-date and accurate data available for the homeless population on Opendatatoronto, it may still have a bias toward the estimation of the true homeless population in the city of Toronto.
-For example, people would not stay in the shelter every day and they would enter and exit the shelter frequently [@toronto_shelter_data].
-Accurately measuring the true population demographic is a very difficult topic even in Canada, but the shelter flow data would provide a proper estimate of our research interest [@dionne2023review].
-To understand the problems better and develop more strategies, the next steps would be to research sampling methods that could reach out and survey the population that is difficult to contact such as the unsheltered homeless population.
-More incorporative sampling strategies would provide us better understanding of homeless demographics and develop further sophisticated plans to reduce homelessness.
+Although our dataset is the most up-to-date and accurate information available on Opendatatoronto, it remains biased in estimating the true homeless population in Toronto.
+Individuals do not stay in shelters continuously and frequently cycle in and out of the system [@toronto_shelter_data].
+Additionally, as noted in the data section, an estimated 18% of the homeless population does not use the shelter system, making our data underrepresentative of this group.
+Accurately capturing the full demographic of the homeless population is challenging, even in Canada, but the shelter flow data provides a reasonable estimate for our research purposes [@dionne2023review].
+To improve understanding and develop more effective strategies, future research should explore sampling methods that reach hard-to-contact groups, such as the unsheltered homeless.
+More inclusive sampling approaches would offer a clearer picture of homeless demographics and enable more targeted interventions to reduce homelessness.
 

### 2024-09-27 paper/paper.qmd
final edition

@@
 While the city has made progress in expanding housing, we find that a housing-centric approach may not be sufficient to address the needs of the evolving homeless population.
-Our analysis reveals significant demographic changes in Toronto’s homeless population over the past six years, including increases in the proportion of chronically homeless individuals, refugees, and working-age adults.
-These demographic shifts suggest that the traditional emphasis on housing development alone is no longer be an adequate solution.
+Our analysis find significant demographic changes in Toronto’s homeless population over the past six years, including increases in the proportion of chronically homeless individuals, refugees, and working-age adults.
+These demographic shifts indicate that the traditional emphasis on housing development alone is no longer be an adequate solution.
 
 In this paper, we explore these demographic changes using monthly data from Toronto's Shelter Management Information System (SMIS) from January 2018 to June 2024 [@toronto_shelter_data].
-We detail the data collection process, cleaning methods, and key findings in the Data section.
-In the final section, Discussion, we assess the implications of our results for policy and propose recommendations for more comprehensive strategies that extend beyond housing solutions to address the diverse needs of Toronto's homeless population.
+We detail the data collection process, cleaning methods, and key findings in the **Data** section.
+In the final section, **Discussion**, we assess the implications of our results for policy and propose recommendations for more comprehensive strategies that extend beyond housing solutions to address the diverse needs of Toronto's homeless population.
 
@@
 
-While other datasets, such as daily shelter occupancy, are available, they lack the demographic depth necessary for this study.
+While other datasets, such as daily shelter occupancy, are available, they lack the demographic information necessary for this study.
 Data loading, cleaning, and analysis were carried out in **R** [@citeR], with packages **tidyverse** [@tidy], **here** [@here], **styler** [@style], and **knitr** [@knitr].
-This comprehensive approach ensures that our analysis accurately reflects the realities faced by individuals experiencing homelessness in Toronto.
 
@@
 Here, according to [@chronic], chronic homelessness refers to "People who meet one of the two following criteria, as per the federal definition of chronic homelessness.
-The person has recorded a minimum of 180 overnight stays in the past year (365 days), or the person has recurrent overnight stays over the past three years with a cumulative duration of at least 546 nights." We can see from the graph that the proportion of chronic homeless increased a lot, from less than 35% in the first quarter of 2018 to more than 65% in the second quarter of 2024.
+The person has recorded a minimum of 180 overnight stays in the past year (365 days), or the person has recurrent overnight stays over the past three years with a cumulative duration of at least 546 nights." We can see from the graph that the proportion of chronic homeless nealy doubled, from less than 35% in the first quarter of 2018 to more than 65% in the second quarter of 2024.
 ```{r}
@@
 @fig-WorkingAge displays the trend of the proportion of different working age statuses among the actively homeless population over the quarters of each year.
-Here, the working age statuses are categorized into three groups: Early Career Stage, Working Age, and Pre-Retirement and Retirement Age.
 Even though the trend is not dramatic, We can see from the graph that the proportion of the Early Career Stage population as well as pre-retire and retired has decreased (from 29% and 17% separately in the first quarter of 2018 to 24.5% and 14.5% in the second quarter in 2024).
@@
 
-First, as the result section points out, the proportion of the chronically homeless population nearly doubled in the past six years.
-Despite its nature of long-lasting homeless status, studies have also shown that chronic homelessness is closely associated with chronic illness or other mental issues @chronic.
-Research further indicates that the average income of people experiencing mental illness and different types of disorders is significantly lower than normal people's income @Sareen2011.
+First, as the result section points out, the proportion of the chronically homeless population increased 30% in the past six years.
+Despite its nature of long-lasting homeless status, studies have also shown that chronic homelessness is closely associated with chronic illness or other mental issues [@chronic].
+Research further indicates that the average income of people experiencing mental illness and different types of disorders is significantly lower than normal people's income [@Sareen2011].
 This makes it particularly difficult for people who have already experienced health problems to escape homelessness.

### 2024-09-27 scripts/00-simulate_data.R
small update

@@ -65,7 +65,7 @@ population_data <- data.frame(
   "Retirement" = retirement
 )
 
-write_csv(population_data, "data/raw_data/simulated_data.csv")
+write_csv(population_data, "data/simulated_data/simulated_data.csv")
 
 
 # Tests

### 2024-09-27 scripts/04-test_simulated_data.R

@@ -1,0 +1,75 @@
+#### Preamble ####
+# Purpose: Test simulated homeless population data
+# Author: Junbo Li
+# Date: SEP-20-2024
+# Contact: junb.li@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+
+
+#### Test data ####
+data <- read.csv("data/simulated_data/simulated_data.csv")
+
+# Test 1: Ensure all cells are non-empty
+# Function to check if any cell in the data frame is empty (NA or "")
+check_empty_cells <- function(df) {
+  # Check for NA values
+  has_na <- any(is.na(df))
+  
+  # Check for empty strings in character columns
+  has_empty_strings <- any(sapply(df, function(x) {
+    if (is.character(x)) {
+      any(x == "")
+    } else {
+      FALSE
+    }
+  }))
+  
+  # Return TRUE if either NA or empty strings are found
+  return(has_na || has_empty_strings)
+}
+
+# Apply the function
+if (check_empty_cells(population_data)) {
+  print("The data frame contains empty cells (NA or empty strings).")
+} else {
+  print("The data frame has no empty cells.")
+}
+
+
+
+# Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
+test_sum_equals_total <- all(population_data$Pre_Working +
+                               population_data$Working +
+                               population_data$Retirement == population_data$actively_homeless)
+
+if (test_sum_equals_total) {
+  print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
+} else {
+  print("Test 2 Failed: The sum of the working categories does not equal the total population.")
+}
+
+
+# Test 3: Ensure all numerical cells are greater than 0
+# Function to check if all numerical columns have values greater than 0
+test_numerical_gt_zero <- function(df) {
+  # Apply the condition only to numeric columns
+  numeric_columns <- sapply(df, is.numeric)
+  all(sapply(df[, numeric_columns], function(x) all(x > 0)))
+}
+
+# Apply the test
+result <- test_numerical_gt_zero(population_data)
+
+# Display the result
+if (result) {
+  print("All numerical inputs are greater than 0")
+} else {
+  print("Some numerical inputs are less than or equal to 0")
+}

### 2024-09-27 scripts/00-simulate_data.R
delete test in simulation

@@ -67,63 +67,3 @@ population_data <- data.frame(
 
 write_csv(population_data, "data/simulated_data/simulated_data.csv")
 
-
-# Tests
-
-# Test 1: Ensure all cells are non-empty
-# Function to check if any cell in the data frame is empty (NA or "")
-check_empty_cells <- function(df) {
-  # Check for NA values
-  has_na <- any(is.na(df))
-
-  # Check for empty strings in character columns
-  has_empty_strings <- any(sapply(df, function(x) {
-    if (is.character(x)) {
-      any(x == "")
-    } else {
-      FALSE
-    }
-  }))
-
-  # Return TRUE if either NA or empty strings are found
-  return(has_na || has_empty_strings)
-}
-
-# Apply the function
-if (check_empty_cells(population_data)) {
-  print("The data frame contains empty cells (NA or empty strings).")
-} else {
-  print("The data frame has no empty cells.")
-}
-
-
-
-# Test 2: Ensure the sum of Pre-working, Working, and Retirement equals the Total Population
-test_sum_equals_total <- all(population_data$Pre_Working +
-  population_data$Working +
-  population_data$Retirement == population_data$actively_homeless)
-
-if (test_sum_equals_total) {
-  print("Test 2 Passed: The sum of Pre-working, Working, and Retirement equals the Total Population.")
-} else {
-  print("Test 2 Failed: The sum of the working categories does not equal the total population.")
-}
-
-
-# Test 3: Ensure all numerical cells are greater than 0
-# Function to check if all numerical columns have values greater than 0
-test_numerical_gt_zero <- function(df) {
-  # Apply the condition only to numeric columns
-  numeric_columns <- sapply(df, is.numeric)
-  all(sapply(df[, numeric_columns], function(x) all(x > 0)))
-}
-
-# Apply the test
-result <- test_numerical_gt_zero(population_data)
-
-# Display the result
-if (result) {
-  print("All numerical inputs are greater than 0")
-} else {
-  print("Some numerical inputs are less than or equal to 0")
-}
